---
title: "Module 2 — Intro to R and Quarto"
format: html
editor: source
execute:
  warning: false
  message: false
---

This is your workspace for the Module 2 worksheet. The full instructions live
on the lesson page — keep it open alongside this notebook. You will build up
this notebook one part at a time: fill in the empty chunk under each heading,
press **Run Cell** to see what happens, and write a sentence of your own
between the chunks. Render with **Preview** whenever the worksheet tells you to.

Always load the packages you need in a setup chunk at the top. Run this first,
every time you open the notebook:

```{r setup}
library(tidyverse)
library(here)
```

## Part 1 — Run a chunk, write a sentence, render

```{r}
# your code here
```

## Part 2 — Read the data and look at it

```{r}
# your code here
```

## Part 3 — Pull out columns and summarise them

```{r}
# your code here
```

## Part 4 — Build something with Pi

```{r}
# your code here
```

## Capstone — make it your own

```{r}
# your code here
```

## Before you go

Pick one step today that confused you. Ask Pi to explain it, then push back:
*"why did you use X instead of Y?"* Notice when Pi sounds confident versus when
it is guessing.
