---
title: "Module 3 — Data visualization"
format: html
editor: source
execute:
  warning: false
  message: false
---

> *Mode flip.* If you were **agentic** in Exercise 1, you are **manual** now; if you were manual, you are agentic.

```{r setup}
library(tidyverse)
library(here)

table_02 <- read_csv(here("data/instructional_dataset/02_visit_clinical_measurements_UKZN_workshop_2023.csv"))
```

## Part 1 — Read two plots (no code)

```{r}
#| echo: false
#| fig-width: 6
#| fig-height: 4
ggplot(table_02, aes(x = ph, y = nugent_score, color = arm)) +
  geom_jitter()
```

**Plot 1 (above). Answer in plain words:**

1. What is the mark?
2. What mark property comes from what column?
3. How many rows of data made each mark?

```{r}
#| echo: false
#| fig-width: 6
#| fig-height: 4
ggplot(table_02, aes(x = arm, y = ph, fill = arm)) +
  geom_boxplot()
```

**Plot 2 (above). Answer the same three questions.**

**Sketch.** Draw on paper or describe in words a plot that answers "does pH differ between treatment arms?" Name the mark, the mappings, and how many rows make each mark.

*Your sketch / description:*

## Part 2 — Build the plot

Each step adds one piece on top of the previous plot. Run after each.

```{r}
# step 1: ggplot(table_02, aes(x = ph, y = nugent_score)) + geom_point()
```

```{r}
# step 2: swap geom_point() for geom_jitter()
```

```{r}
# step 3: add aes(color = arm)
```

```{r}
# step 4: switch to a boxplot of ph by arm: aes(x = arm, y = ph) + geom_boxplot()
```

```{r}
# step 5a: aes(fill = arm) — mapping
```

```{r}
# step 5b: fill = "navyblue" outside aes() — parameter
```

```{r}
# step 6: add labs(x = "Treatment arm", y = "Vaginal pH", fill = "Arm")
```

```{r}
# step 7: add facet_wrap(~ time_point)
```

## Part 3 — Reach (optional)

Produce a polished version of the final boxplot:

- faceted by `time_point`
- colored by `arm`
- axes and legend labeled
- panels in chronological order (baseline, week_1, week_7)
- a legible theme

```{r}
# reach: the polished final figure
```
