---
title: "Module 4 — Visualizing distributions and summaries"
format: html
editor: source
execute:
  warning: false
  message: false
---

> *Mode flip.* If you were **agentic** in Exercise 1, you are **manual** now; if you were manual, you are agentic.

```{r setup}
library(tidyverse)
library(here)
library(ggbeeswarm)

table_02 <- read_csv(here("data/instructional_dataset/02_visit_clinical_measurements_UKZN_workshop_2023.csv"))
```

## Part 1 — Read three plots (no code)

```{r}
#| echo: false
#| fig-width: 6
#| fig-height: 4
ggplot(table_02, aes(ph)) +
  geom_histogram(binwidth = 0.2, fill = "navyblue") +
  facet_wrap(~ arm)
```

**Plot 1 (above). Answer in plain words:**

1. What is the mark?
2. What mark property comes from what column?
3. How many rows of data made each mark?

```{r}
#| echo: false
#| fig-width: 6
#| fig-height: 4
ggplot(table_02, aes(arm, ph)) +
  geom_boxplot()
```

**Plot 2 (above). Answer the same three questions.**

```{r}
#| echo: false
#| fig-width: 6
#| fig-height: 4
ggplot(table_02, aes(arm, ph)) +
  geom_quasirandom()
```

**Plot 3 (above). Answer the same three questions.**

**Sketch.** Draw on paper or describe in words a plot that would answer "does pH differ between arms?" using a **layered** approach — raw observations underneath a summary mark.

*Your sketch / description:*

## Part 2 — Build the plot

```{r}
# step 1: ggplot(table_02, aes(ph)) + geom_histogram(), then add + facet_wrap(~ arm)
```

```{r}
# step 2: ggplot(table_02, aes(arm, ph)) + geom_boxplot()
```

```{r}
# step 3: swap geom_boxplot() for geom_quasirandom()
```

```{r}
# step 4: swap for geom_violin()
```

```{r}
# step 5a: compute ph_summary <- table_02 |> group_by(arm) |>
#                                summarise(mean_ph = mean(ph), sd_ph = sd(ph))
ph_summary <- NULL  # replace NULL with the pipeline
ph_summary
```

```{r}
# step 5b: ggplot(ph_summary, aes(arm, mean_ph)) +
#            geom_pointrange(aes(ymin = mean_ph - sd_ph, ymax = mean_ph + sd_ph))
```

```{r}
# step 6: layered — geom_quasirandom(data = table_02, aes(arm, ph), alpha = 0.5)
#                 + geom_pointrange(data = ph_summary, aes(arm, mean_ph,
#                                                          ymin = mean_ph - sd_ph,
#                                                          ymax = mean_ph + sd_ph),
#                                   color = "red", size = 0.8)
```

```{r}
# step 7: faceted from long_table (the pivot_longer result from Exercise 1)
#         long_table |> ggplot(aes(arm, value)) + geom_boxplot() +
#                       facet_wrap(~ measurement, scales = "free_y")
# If you don't have long_table available, rebuild it:
#   long_table <- table_02 |> pivot_longer(c(ph, nugent_score, crp_blood),
#                                          names_to = "measurement",
#                                          values_to = "value")
```

## Part 3 — Reach (optional)

Build the polished final figure: the layered plot from step 6, faceted by `time_point` (in chronological order: `baseline`, `week_1`, `week_7`), with axis labels and a clean theme.

```{r}
# reach: the polished final figure
```

Optional add-on: density overlay on the histogram from step 1.

```{r}
# reach add-on:
# ggplot(table_02, aes(ph)) +
#   geom_histogram(aes(y = after_stat(density)), binwidth = 0.2, fill = "navyblue", alpha = 0.5) +
#   geom_density(color = "navyblue") +
#   facet_wrap(~ arm)
```
