---
title: "Module 4 — Reshaping and combining tables"
format: html
editor: source
execute:
  warning: false
  message: false
---

> *Your mode for this exercise.* Group A: **agentic** — describe each step to Pi and let it write the code. Group B: **manual** — write each step yourself. A chat LLM may answer conceptual questions but should not edit this notebook or run commands. Exercise 2 flips the modes.

```{r setup}
library(tidyverse)
library(here)

table_01 <- read_csv(here("data/instructional_dataset/01_participant_metadata_UKZN_workshop_2023.csv"))
table_02 <- read_csv(here("data/instructional_dataset/02_visit_clinical_measurements_UKZN_workshop_2023.csv"))
```

## Part 1 — Predict (no code)

For each pipeline below, write in plain words *how many rows* the result has and *what one row represents*. Don't run any code yet.

```{r}
table_02
```

**1. After `table_02 |> group_by(pid) |> summarise(mean_ph = mean(ph))`:**

*Your answer:*

**2. After `table_02 |> group_by(arm, time_point) |> summarise(mean_ph = mean(ph))`:**

*Your answer:*

**3. After `per_person |> left_join(table_01, by = "pid")`, where `per_person` is the 44-row result from prediction 1:**

*Your answer:*

**4. After `arm_timepoint |> pivot_wider(names_from = time_point, values_from = mean_ph)`, where `arm_timepoint` is the 6-row result from prediction 2:**

*Your answer:*

## Part 2 — Build the pipeline

```{r}
# step 1: pipe table_02 into head() to see the starting shape
```

```{r}
# step 2: group_by(pid) |> summarise(mean_ph = mean(ph), mean_nugent = mean(nugent_score))
# assign it to per_person so step 3 can use it
per_person <- NULL  # replace NULL with the pipeline
per_person
```

```{r}
# step 3: per_person |> left_join(table_01, by = "pid")
```

```{r}
# step 4: separately, table_02 |> group_by(arm, time_point) |> summarise(mean_ph = mean(ph))
# assign it to arm_timepoint
arm_timepoint <- NULL  # replace NULL with the pipeline
arm_timepoint
```

```{r}
# step 5: arm_timepoint |> pivot_wider(names_from = time_point, values_from = mean_ph)
```

```{r}
# step 6: separately, table_02 |> pivot_longer(c(ph, nugent_score, crp_blood),
#                                              names_to = "measurement", values_to = "value")
# assign it to long_table — Exercise 2 will use this
long_table <- NULL  # replace NULL with the pipeline
long_table
```

## Part 3 — Reach (optional)

```{r}
# reach: smokers <- table_01 |> filter(smoker == "smoker")
# then: per_person |> semi_join(smokers, by = "pid")
# then: group_by(arm) |> summarise(...) for per-arm pH means among smokers only
```
