---
title: "Module 6 — Figures with p-values"
format: html
editor: source
execute:
  warning: false
  message: false
---

> *Mode flip.* If you were **agentic** in Exercise 1, you are **manual** now; if you were manual, you are agentic.

```{r setup}
library(tidyverse)
library(here)
library(gtsummary)
library(ggpubr)

table_00 <- read_csv(here("data/instructional_dataset/00_sample_ids_UKZN_workshop_2023.csv"))
table_01 <- read_csv(
  here("data/instructional_dataset/01_participant_metadata_UKZN_workshop_2023.csv"),
  col_types = cols(sex = col_character())
)
table_02 <- read_csv(here("data/instructional_dataset/02_visit_clinical_measurements_UKZN_workshop_2023.csv"))
table_03 <- read_csv(here("data/instructional_dataset/03_elisa_cytokines_UKZN_workshop_2023.csv"))
```

## Part 1 — Read three plots (no code)

```{r}
#| echo: false
#| fig-width: 6
#| fig-height: 4
table_02 |>
  filter(time_point == "week_7") |>
  ggboxplot(x = "arm", y = "ph", fill = "arm") +
  stat_compare_means(method = "wilcox.test", label.x = 1.3)
```

**Plot 1 (above). Answer in plain words:**

1. What is the mark?
2. What is the comparison being tested?
3. What test is on it?
4. How many rows of data made each mark?

```{r}
#| echo: false
#| fig-width: 9
#| fig-height: 6
cytokines_keep <- c("IL-1a", "IL-6", "IL-8", "IL-1b", "MIG", "IP-10")
cytokine_panel_df <- table_03 |>
  filter(cytokine %in% cytokines_keep, limits == "within limits") |>
  left_join(table_00, by = "sample_id") |>
  filter(time_point == "week_7")

ggplot(cytokine_panel_df, aes(arm, conc, fill = arm)) +
  geom_boxplot() +
  facet_wrap(~ cytokine, scales = "free_y") +
  stat_compare_means(method = "wilcox.test", label.x = 1.3) +
  theme(legend.position = "none")
```

**Plot 2 (above). Answer the same four questions for each panel.**

```{r}
#| echo: false
#| fig-width: 8
#| fig-height: 4
pre_post_df <- table_02 |>
  filter(time_point %in% c("baseline", "week_7")) |>
  mutate(time_point = fct_relevel(time_point, "baseline", "week_7"))

ggplot(pre_post_df, aes(time_point, ph)) +
  geom_line(aes(group = pid), alpha = 0.4) +
  geom_point(aes(color = arm), size = 2) +
  facet_wrap(~ arm) +
  stat_compare_means(
    method = "wilcox.test",
    paired = TRUE,
    comparisons = list(c("baseline", "week_7"))
  )
```

**Plot 3 (above). Answer the same four questions.**

## Part 2 — Build

```{r}
# step 1: ggboxplot of ph by arm at week_7 + stat_compare_means(method = "wilcox.test")
```

```{r}
# step 2: same plot in plain ggplot (geom_boxplot + stat_compare_means)
# confirm both give the same p-value
```

```{r}
# step 3: the cytokine panel
# filter table_03 to 6 cytokines and within-limits, left_join table_00,
# filter to week_7, then ggplot + geom_boxplot + facet_wrap + stat_compare_means
```

```{r}
# step 4: paired pre-post
# filter table_02 to baseline + week_7, mutate(time_point = fct_relevel(...)),
# then ggplot + geom_line(aes(group = pid)) + geom_point + facet_wrap(~ arm) +
#   stat_compare_means(method = "wilcox.test", paired = TRUE,
#                      comparisons = list(c("baseline", "week_7")))
```

```{r}
# step 5: adjusted analysis
# outcome_table <- table_02 |> filter(time_point == "week_7") |>
#                  left_join(table_01, by = c("pid", "arm"))
# lm(ph ~ arm + age + smoker, data = outcome_table) |> tbl_regression()
```

## Part 3 — Reach (optional)

Pick *one* analysis above and polish it for publication, or repeat the full arc on `nugent_score` or `crp_blood`.

```{r}
# reach: polished version or alternative outcome
```
