---
title: "Module 6 — Demographics and Table 1"
format: html
editor: source
execute:
  warning: false
  message: false
---

> *Your mode for this exercise.* Group A: **agentic** — describe each step to Pi and let it write the code. Group B: **manual** — write each step yourself. A chat LLM may answer conceptual questions but should not edit this notebook or run commands. Exercise 2 flips the modes.

```{r setup}
library(tidyverse)
library(here)
library(gtsummary)

table_01 <- read_csv(
  here("data/instructional_dataset/01_participant_metadata_UKZN_workshop_2023.csv"),
  col_types = cols(sex = col_character())
)
```

## Part 1 — Predict (no code)

```{r}
table_01
```

For each non-pid column of `table_01` (`arm`, `smoker`, `age`, `education`, `sex`), predict in plain words below. Don't run any code yet.

**1. `arm`**

- How will `tbl_summary(by = arm)` summarise it? *Your answer:*
- What test will `add_p()` pick? *Your answer:*

**2. `smoker`**

- *Your answer (summary, test):*

**3. `age`**

- *Your answer (summary, test):*

**4. `education`**

- *Your answer (summary, test):*

**5. `sex`**

- *Your answer (summary, test):*

## Part 2 — Build

Each step adds one piece on top of the previous table. Run after each step.

```{r}
# step 1: tbl_summary defaults
table_01 |>
  select(-pid, -sex) |>
  tbl_summary(by = arm)
```

```{r}
# step 2: add p-values
# (modify the previous chunk by adding `|> add_p()`)
```

```{r}
# step 3: override age to mean (SD) + t-test
# statistic = list(age ~ "{mean} ({sd})")
# test = list(age ~ "t.test")
```

```{r}
# step 4: polish
# add `|> add_overall() |> bold_labels() |> modify_header(label ~ "**Characteristic**")`
```

## Part 3 — Reach (optional)

```{r}
# reach: same table, but by = smoker instead of by = arm
# what question is this table now answering?
```
