                                                                    ---
title: "Module 7 — Yogurt Study Demographics (Table 1)"
format: html
editor: source
execute:
  warning: false
  message: false
---

```{r setup}
library(tidyverse)
library(here)
library(gtsummary)

# Participant metadata — one row per person
table_01 <- read_csv(
  here("data/group_a_yogurt/01_participant_metadata_yogurt.csv"),
  col_types = cols(sex = col_character())
)
```

## Preview the data

```{r}
table_01
```

Columns: `pid` (participant ID), `arm` (yogurt / unchanged_diet),
`days_since_last_sex`, `birth_control`, `age`, `education`, `sex`.

There are **51 participants** split between the two study arms.

---

## Basic Table 1 — by study arm

```{r}
# step 1: default tbl_summary grouped by arm
table_01 |>
  select(-pid) |>
  tbl_summary(by = arm)
```

```{r}
# step 2: add p-values
table_01 |>
  select(-pid) |>
  tbl_summary(by = arm) |>
  add_p()
```

```{r}
# step 3: polished — overall column + bold labels + nicer header
table_01 |>
  select(-pid) |>
  tbl_summary(by = arm) |>
  add_p() |>
  add_overall() |>
  bold_labels() |>
  modify_header(label ~ "**Characteristic**")
```

---

## Part 1 — Predict (no code)

For each column below, predict in plain words:

**1. `arm`**
- How will `tbl_summary(by = arm)` summarise it?
- What test will `add_p()` pick?

**2. `days_since_last_sex`**
- *Your answer (summary, test):*

**3. `birth_control`**
- *Your answer (summary, test):*

**4. `age`**
- *Your answer (summary, test):*

**5. `education`**
- *Your answer (summary, test):*

**6. `sex`**
- *Your answer (summary, test):*

---

## Part 2 — Explorations (optional)

Try changing `by = arm` to `by = birth_control` and see what question the table answers.

```{r}
# reach: table grouped by birth_control instead of arm
```
