---
title: "Module 8 — Yogurt Study Boxplots with faceting"
format: html
editor: source
execute:
  warning: false
  message: false
---

```{r setup}
library(tidyverse)
library(here)
library(ggpubr)

# Participant metadata
table_01 <- read_csv(
  here("data/group_a_yogurt/01_participant_metadata_yogurt.csv"),
  col_types = cols(sex = col_character())
)
```

## Preview

```{r}
table_01
```

---

## Basic boxplot — age by study arm

```{r}
ggboxplot(table_01, x = "arm", y = "age")
```

## Add pairwise comparison

```{r}
ggboxplot(table_01, x = "arm", y = "age") +
  stat_compare_means()
```

The default test is Wilcoxon rank-sum test (Mann–Whitney U) for 2 groups.

---

## With points + color

```{r}
ggboxplot(table_01, x = "arm", y = "age",
          #add = "jitter",
          fill="arm", palette = c("#E7B800", "#00AFBB")) +
  stat_compare_means()
```

---

## Facet by education

Now let's see the same plot split by education level using `facet_wrap()`.

```{r}
ggboxplot(table_01, x = "arm", y = "age",
          add = "jitter",
          color = "arm", palette = c("#E7B800", "#00AFBB")) +
  stat_compare_means() +
  facet_wrap(~ education)
```

Here `facet_wrap(~ education)` creates one small panel for each education
level, so you can see whether the age pattern differs across education groups.

---

## Facet by birth control

```{r}
ggboxplot(table_01, x = "arm", y = "age",
          add = "jitter",
          color = "arm", palette = c("#E7B800", "#00AFBB")) +
  stat_compare_means() +
  facet_wrap(~ birth_control)
```

---

## Days since last sex — faceted by education

```{r}
ggboxplot(table_01, x = "arm", y = "days_since_last_sex",
          add = "jitter",
          color = "arm", palette = c("#E7B800", "#00AFBB")) +
  stat_compare_means() +
  facet_wrap(~ education)
```

---

## Two-way faceting: education × birth control

You can also facet by **two variables** with `facet_grid()`:

```{r}
ggboxplot(table_01, x = "arm", y = "age",
          add = "jitter",
          color = "arm", palette = c("#E7B800", "#00AFBB")) +
  stat_compare_means() +
  facet_grid(birth_control ~ education)
```

`facet_grid(birth_control ~ education)` makes a grid where rows are
birth control types and columns are education levels.

---

## Try it yourself

Try changing `y` to another numeric column, or change the faceting variable!

```{r}
# experiment here
```