---
title: "Module 9 — Mixed Model (lmer) + Regression Table (Yogurt Study)"
format: html
editor: source
execute:
  warning: false
  message: false
---

> **Your mode for this exercise.** Group A: **agentic** — describe each step to Pi
> and let it write the code. Group B: **manual** — write each step yourself.

```{r setup}
library(tidyverse)
library(here)
library(lme4)       # for lmer() — linear mixed-effects models
library(gtsummary)  # for tbl_regression()
library(broom.mixed) # helps gtsummary understand lmer models
```

The yogurt study has **two time points per participant** (baseline and after
antibiotic), so each person contributes 2 rows of data. We use a **mixed
model** (`lmer`) to handle this — `(1 | pid)` means "each person gets their
own random intercept."

```{r}
# Load and join the data
samples <- read_csv(here("data/group_a_yogurt/00_sample_ids_yogurt.csv"))
qpcr    <- read_csv(here("data/group_a_yogurt/02_qpcr_results_yogurt.csv"))
meta <- read_csv(
  here("data/group_a_yogurt/01_participant_metadata_yogurt.csv"),
  col_types = cols(sex = col_character())
)

# Join qPCR results → sample mapping → participant metadata
df <- samples |>
  left_join(qpcr, by = "sample_id") |>
  left_join(meta, by = c("pid", "arm")) |>
  mutate(
    # Make factors so R knows the reference groups
    time_point = factor(time_point, levels = c("baseline", "after_antibiotic")),
    arm        = factor(arm,        levels = c("unchanged_diet", "yogurt")),
    # Log-transform the bacterial counts (they're huge numbers)
    log_bacteria = log(qpcr_bacteria)
  )
```

---

## Preview the data

```{r}
df |>
  select(pid, time_point, arm, qpcr_bacteria, log_bacteria) |>
  head(10)
```

We have 51 participants × 2 time points = 102 rows.

---

## Part 1 — Predict (no code)

Before running the model, answer in plain words:

1. The model formula is `log_bacteria ~ time_point * arm + (1 | pid)`.
   What does `(1 | pid)` mean? *Your answer:*

2. Which coefficient do you expect to be **largest** (biggest effect)?
   *Your answer:*

3. The interaction term `time_point * arm` asks: "Does the change from
   baseline to after-antibiotic differ between the yogurt and unchanged-diet
   groups?" If yogurt helps protect bacteria during antibiotics, what sign
   would the interaction coefficient have (positive or negative)?
   *Your answer:*

---

## Part 2 — Build

Each step adds one piece. Run after each step.

```{r}
# step 1: fit the mixed model
m <- lmer(log_bacteria ~ time_point * arm + (1 | pid), data = df)
```

```{r}
# step 2: examine the model summary
summary(m)
```

Look at the **Fixed effects** table. The key numbers:

| Coefficient | Estimate | Interpretation |
|---|---|---|
| (Intercept) | `r round(fixef(m)[1], 2)` | Baseline log(bacteria) in unchanged_diet group |
| after_antibiotic | `r round(fixef(m)[2], 2)` | Change after antibiotic (unchanged_diet) |
| yogurt | `r round(fixef(m)[3], 2)` | Difference at baseline (yogurt vs unchanged) |
| after_antibiotic × yogurt | `r round(fixef(m)[4], 2)` | Extra change in yogurt group |

```{r}
# step 3: display the model as a publication-ready regression table
tbl_regression(m)
```

Notice `tbl_regression()` automatically shows:
- The **Beta** column (coefficient estimates)
- **95% confidence intervals** (using profile likelihood)

```{r}
# step 4: add overall p-values for multi-level terms
tbl_regression(m) |>
  add_global_p()
```

`add_global_p()` tests "is this term significant overall?" — for the
interaction, it asks "does the time-course differ between arms?"

---

## Part 3 — Explore (optional)

Try changing the response variable to one of the specific bacterial species:

```{r}
# Instead of total bacteria, model L. crispatus
df <- df |>
  mutate(log_crispatus = log(qpcr_crispatus + 1))  # +1 avoids log(0)

m_crispatus <- lmer(log_crispatus ~ time_point * arm + (1 | pid), data = df)
tbl_regression(m_crispatus) |>
  add_global_p()
```

```{r}
# Or model L. iners
df <- df |>
  mutate(log_iners = log(qpcr_iners + 1))

m_iners <- lmer(log_iners ~ time_point * arm + (1 | pid), data = df)
tbl_regression(m_iners) |>
  add_global_p()
```

```{r}
# Or add age as a covariate
m_with_age <- lmer(log_bacteria ~ time_point * arm + age + (1 | pid), data = df)
tbl_regression(m_with_age) |>
  add_global_p()
```

---

## What We Learned

- **`lmer()`** from the `lme4` package fits linear mixed-effects models for
  repeated measures data
- **`(1 | pid)`** adds a random intercept per participant, accounting for
  the fact that each person has their own baseline level
- **`tbl_regression()`** from `gtsummary` turns the model output into a
  publication-ready table with coefficients and confidence intervals
- **`add_global_p()`** adds overall p-values for each term in the model
