---
title: "Module 10 — Yogurt Study: Combining Data + Summary Table"
format: html
editor: source
execute:
  warning: false
  message: false
---

> **Your mode for this exercise.** Group A: **agentic** — describe each step to Pi
> and let it write the code. Group B: **manual** — write each step yourself.

```{r setup}
library(tidyverse)
library(here)
library(gtsummary)
```

We have two CSV files about the yogurt study:

- **`00_sample_ids_yogurt.csv`** — one row per *sample* (each participant has
  a baseline sample and an after-antibiotic sample)
- **`01_participant_metadata_yogurt.csv`** — one row per *participant* (age,
  education, birth control, etc.)

We'll join them together, then make a summary table grouped by study arm.

---

## Step 1 — Load both files

```{r}
samples  <- read_csv(here("data/group_a_yogurt/00_sample_ids_yogurt.csv"))
metadata <- read_csv(
  here("data/group_a_yogurt/01_participant_metadata_yogurt.csv"),
  col_types = cols(sex = col_character())
)
```

Take a quick look at each:

```{r}
samples
```

```{r}
metadata
```

---

## Step 2 — Join them

`left_join()` matches rows by the column `pid`. Each sample gets the
participant's age, education, etc. attached to it.

```{r}
df <- samples |> 
  left_join(metadata, by = "pid")

df
```

Now every row has both the sample info *and* the participant's metadata.

---

## Step 3 — Summary table by study arm

Because we joined the data, we can group the 102 samples by `arm`
and see whether the two groups look similar across different variables.

```{r}
# step 1: default tbl_summary grouped by arm
df |> 
  select(arm, age, education, birth_control, days_since_last_sex) |> 
  tbl_summary(by = arm)
```

```{r}
# step 2: add p-values
df |> 
  select(arm, age, education, birth_control, days_since_last_sex) |> 
  tbl_summary(by = arm) |> 
  add_p()
```

```{r}
# step 3: polished — overall column + bold labels + nicer header
df |> 
  select(arm, age, education, birth_control, days_since_last_sex) |> 
  tbl_summary(by = arm) |> 
  add_p() |> 
  add_overall() |> 
  bold_labels() |> 
  modify_header(label ~ "**Characteristic**")
```

---

## Part 1 — Predict (no code)

Before running the code above, predict in plain words what will happen:

**1. `age`**
- How will `tbl_summary(by = arm)` summarise it?
- What test will `add_p()` pick?

**2. `education`**
- How will `tbl_summary(by = arm)` summarise it?
- What test will `add_p()` pick?

**3. `birth_control`**
- Same two questions:

**4. `days_since_last_sex`**
- Same two questions:

---

## Part 2 — Explorations (optional)

Try changing `by = arm` to `by = time_point` and see what changes.

```{r}
# reach: table grouped by time_point (baseline vs after_antibiotic)
```

Or try grouping by `birth_control` instead:

```{r}
# reach: table grouped by birth_control method
```

What question does each version of the table answer?